Still need a good title

### 2010-07-03 pkg/paper/Draft/draft-Z-W.tex
smaller head at draft
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/glogis@25 b679b2f7-6539-45d0-a605-5ddcb976cde9
- Simplified header further. - Reduced spacing and increased textwidth (via a4wide), hence equations   don't need special split treatment anymore. - Put table into a regular {table}/{tabular} environment. Changed formatting   somewhat. - Changed citation style (due to problems with the Harvard style) to   abbrvnat, still can choose something better later on. - Still many problems with citations: some missing, some BibTeX items   are clearly malformatted.
git-svn-id: svn://sv

\documentclass[titlepage,a4paper, fleqn, flalign]{article} \usepackage[pdftex]{graphicx}
\usepackage[multiple]{footmisc} \pagestyle{myheadings}
\usepackage[latin1]{inputenc}
\usepackage{fancyhdr}
\usepackage{rotating}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage{ltxtable}
\usepackage{tabularx}
\usepackage{rotating}
\usepackage{array}
\usepackage{amssymb}
\usepackage{pstricks}
\usepackage{qtree}
\usepackage{lscape}
\usepackage{setspace}
\usepackage[numbers,round]{natbib}
\usepackage{harvard}

@@

%\title{Draft  \\[2cm]
%\textbf{Structural Breaks in Inflation Rates}}
%
%\author{ \scriptsize by \\
%Mag. Thomas WINDBERGER\\
%Prof. Dr. Achim ZEILEIS \\
%\scriptsize Doctoral Advisors: \\
%\scriptsize Univ.-Prof.Dr. Jesus Crespo-Cuaresma \\
%\scriptsize A.Univ.-Prof.Dr. Janette Walde \\ [1cm]}
%
%
%
%\maketitle

% anderthalb Zeilen Abstand inne

\setstretch{1,5}

\section{Abstract}

The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
%We concentrate on the volatility of the inflation rate rather than its mean since \cite{fried} conjectured that the harmful effect of inflation on growth is driven by inflation volatility.

\newpage
\section{Introduction}

The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2 \%'' \cite{castel}. \cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22). Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank. \\
To further justify the use of inflation volatility, we refer to \cite{rother}: ``Among the harmful effects of inflation, the negative consequences of inflation volatility are of particular concern. High variability of inflation over time makes expectations over the future price level more uncertain. In a world with nominal contracts, this induces risk premia for long-term arrangements, raises costs for hedging against inflation risks and leads to unanticipated redistribution of wealth. Thus, inflation volatility can impede growth even if inflation on average remains restrained.'' \\
The question of interest centers around the way in which a countries decision to join the EMU changed its inflation volatility. There are a number of reasons why the volatility of  one of these nations should change indeed. Given that the country experienced quite volatile inflation rates, its efforts to meet the convergence criterias are likely to lead to a change in the volatility of their respective inflation rates, as has indeed been the case for a number of EMU countries, like Spain, Portugal, Italy and Greece.

@@

Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot ln(HICP_{t}/HICP_{t-1})$. Countries included are Austria,	Belgium,	Czech Republic,	Denmark,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,		Spain,	Sweden,		United Kingdom and the	Euro area.  The data are obtained from the OECD Statistics. \\
The countries can be divided into three different groups: the EMU countries (Portugal, Spain, Italy, Greece, France, Belgium, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.

@@

\newpage
\section{Methodology}

@@

\begin{eqnarray}
& & f(\pi | \theta, \sigma, \delta) = \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
\end{eqnarray}

@@

\begin{eqnarray}
& & E(\pi_i) = \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
& & Var(\pi_i)  = \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
& & Skew(\pi_i) =  \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}

@@

\begin{eqnarray}
& & l(\delta,\theta,\sigma, \pi) = ln(\delta) - ln(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) ln(1+\exp^{-\frac{\pi-\theta}{\sigma}})
\end{eqnarray}

@@

\begin{eqnarray}
& & \psi(\pi_i,\delta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - ln(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
& & \psi(\pi_i,\theta) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})} 
\end{eqnarray}

\begin{equation}
\begin{split}
&& \psi(\pi_i,\sigma) = \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
&& \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
\end{split}
\end{equation}

@@

i.e. first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically as developed in 
%[wann war die Erstpublikation der graphischen Methode?].

%\subsubsection{Kolmogorov-Smirnov Test}
%
%As a goodness of fit test, we use the Kolmogorov Smirnov Test for as given in \cite{massey}.

@@

\newpage
\subsection{Expected Results}

@@

[Estonia should be included]

\begin{longtable}{|l|p{4cm}|p{3cm}|p{3cm}|}
\hline
{\bf Country} & {\bf 1. breakpoint}  & {\bf 2. breakpoint  \\
\hline
Austria (EURO) & Sep. 2007 & none\\
\hline
Belgium (EURO) & Dec. 1999 & none  \\
\hline
Czech Republic &  Jul. 1998  & none \\
\hline
Denmark (ERM~II)& Jun. 2000 & none \\
\hline
Finland (EURO) & none & none  \\
\hline
France (EURO) & Dec. 2004  & none\\
\hline
Germany (EURO) & May 2000 & Dec. 2004 \\
\hline
Greece (EURO) & none & none \\
\hline
Hungary & May 1998 & none\\
\hline
Ireland (EURO) & Mar. 2008 & none\\
\hline
Italy (EURO) & May 1996 &  Dec. 2000  \\
\hline
Luxembourg (EURO) & Dec. 1998 & none\\
\hline
Netherlands  (EURO) & none & none \\
\hline
Poland & May 2001 & none\\
\hline
Portugal (EURO) & Jul. 1992 & Mar. 2004\\
\hline
Slovakia (EURO) & Apr. 1997 & Feb. 2004 \\
\hline
Slovenia (EURO) & Jul. 2003 & none\\
\hline
Spain (EURO) & May 1996  & Dec. 2000\\
\hline
Sweden & Jan. 1993 & none\\
\hline
United Kingdom & Apr. 1992 & none\\
\hline
\end{longtable}

@@

\subsection{Belgium}
%% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
% Country Report Belgium  bei ebscohost in TX all text
There was a  very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000). Volatility  a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate. Economic reason: belgish franc was pegged to German Mark (since 1990). Possibly: Domestic inflation was mainly driven by the surge in world oil prices, with the result that energy prices in Belgium were around 25 \% higher in the first quarter of this year than in the year-earlier period. Also a general upward trend in prices.
% source: Belgium_3206697
tax reductions in personal income tax
% source: Belgium_4611003
inflation turns out to be more volatile in small, open economies such as Belgium (might be the case in some countries, but not in all of them). some explanations happen to be: a) a more volatile currency risk premium (which cannot be the case in Belgium, since it happend after 1.1999 and the currency was pegged thereafter, and even before pegged to Germany); b) degree of exchange--rate--passthrough to inflation (not possible as well) and c) size of monetary policy shocks (if there were any, should have to be observed in ALL EMU countries, since monetary policy equal). another possible explanation are productivity shocks which boost inflation rate volatility (but this for more than 10 years????)

@@

\subsection{Italy}
Drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime. Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy’s decision to take part in the Monetary Union. After 2000 Italy entered an econmic crisis [auf Auffälligkeiten eingehen]
%(http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
%in 1999 italy qualifies for EMU
%strong economic growth from 1996-2001
%2000: economic crisis

@@

\setstretch{1,3}

\newpage
\section{Literature}

\renewcommand*{\refname}{}
\bibliography{papers}
\begin{thebibliography}{99}
\bibliographystyle{apsr}

>>>>

\documentclass[a4paper]{article}
\usepackage{graphicx,a4wide,color,hyperref}
\usepackage[latin1]{inputenc}
\usepackage[authoryear,round]{natbib}
\bibpunct{(}{)}{;}{a}{}{,}
\pagestyle{myheadings}

@@

\title{Still need a good title}
\author{}
\date{}
\maketitle

\begin{abstract}
The aim of this paper is to shed some light on the effect of the European Monetary Union (EMU)  on changing the volatility of the inflation rates in its Member States. 
%We concentrate on the volatility of the inflation rate rather than its mean since \cite{fried} conjectured that the harmful effect of inflation on growth is driven by inflation volatility. 
\end{abstract}

\section{Introduction}

The ECB defines price stability ``as a year-on-year increase in the Harmonised Index of Consumer Prices (HICP) for the euro area of below 2\%'' \cite{castel}. \cite{emerson} makes it clear, that a high inflation rate is also more variable and uncertain and in that way causes more relative price variability, leading to a less efficient price mechanism (p.~22). Thus to achieve a low inflation rate and low variability of inflation must be a key issue for the central bank. \\
To further justify the use of inflation volatility, we refer to \cite{rother}: ``Among the harmful effects of inflation, the negative consequences of inflation volatility are of particular concern. High variability of inflation over time makes expectations over the future price level more uncertain. In a world with nominal contracts, this induces risk premia for long-term arrangements, raises costs for hedging against inflation risks and leads to unanticipated redistribution of wealth. Thus, inflation volatility can impede growth even if inflation on average remains restrained.'' \\
The question of interest centers around the way in which a countries decision to join the EMU changed its inflation volatility. There are a number of reasons why the volatility of  one of these nations should change indeed. Given that the country experienced quite volatile inflation rates, its efforts to meet the convergence criterias are likely to lead to a change in the volatility of their respective inflation rates, as has indeed been the case for a number of EMU countries, like Spain, Portugal, Italy and Greece.

@@

Inflation is measured as the logarithm of the monthly change in the HICP from 1.1990---3.2010, so $x_t = 100\cdot \log(HICP_{t}/HICP_{t-1})$. Countries included are Austria,	Belgium,	Czech Republic,	Denmark,	Finland,	France,	Germany,	Greece,	Hungary,	Ireland,	Italy,	Luxembourg,	Netherlands,	Poland,	Portugal,		Spain,	Sweden,		United Kingdom and the	Euro area.  The data are obtained from the OECD Statistics. \\
The countries can be divided into three different groups: the EMU countries (Portugal, Spain, Italy, Greece, France, Belgium, Netherlands, Germany, Austria, Ireland, Slovenia, Luxembourg and Finland), EU members without ERM~II (Exchange Rate Mechanism): United Kingdom, Sweden, Poland, Czech Republic, Hungary. Denmark stands on its own as a member of the EU and the ERM~II, but not yet a member of the EMU.

@@

\section{Methodology}

@@

\begin{eqnarray}
f(\pi | \theta, \sigma, \delta) & = & \frac{\frac{\delta}{\sigma}*\exp^{-\frac{\pi_i-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi_i-\theta}{\sigma}})^{(\delta+1)}}
\end{eqnarray}

@@

\begin{eqnarray}
E(\pi_i) & = & \theta + \sigma (\gamma(\delta) - \gamma(1)) \\
Var(\pi_i)  & = & \sigma^2(\gamma'(\delta)+\gamma'(1)) \\
Skew(\pi_i) & = & \frac{\gamma''(\delta)-\gamma''(1)}{(\gamma'(\delta)+\gamma'(1))^{3/2}}
\end{eqnarray}

@@

\begin{eqnarray}
l(\delta,\theta,\sigma, \pi) & = & \log(\delta) - \log(\sigma) - \frac{1}{\sigma} (\pi-\theta) - (\delta+1) \log(1+\exp^{-\frac{\pi-\theta}{\sigma}})
\end{eqnarray}

@@

\begin{eqnarray}
\psi(\pi_i,\delta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \delta} = \frac{1}{\delta} - \log(1+\exp^{-\frac{\pi-\theta}{\sigma}}) \\
\psi(\pi_i,\theta) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \theta} = \frac{1}{\sigma} - (\delta+1)*\frac{\frac{1}{\sigma}\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})} 
\end{eqnarray}

\begin{eqnarray}
\psi(\pi_i,\sigma) & = & \frac{\delta l(\delta,\theta,\sigma;\pi)}{\delta \sigma} = -\frac{1}{\sigma} + \frac{1}{\sigma^2}(\pi-\theta) - (\delta+1) * \\
&& \quad \frac{\frac{1}{\sigma^2}(\pi-\theta)\exp^{-\frac{\pi-\theta}{\sigma}}}{(1+\exp^{-\frac{\pi-\theta}{\sigma}})}
\end{eqnarray}

@@

i.e. first the $L_2$ norm is used to aggregate over the components and then the mean of the resulting aggregated process is used as the test statistic. This can also be shown graphically as developed in 
%[wann war die Erstpublikation der graphischen Methode?].
%\subsubsection{Kolmogorov-Smirnov Test}
%% information from: Heiberger/Holland: Statistical Analysis and Data Display
%The one sample test tests whether a random sample is coming from a hypothesized null distribution. It compares the maximum discrepancy between the empirical cumulative distribution of the data, defined as S(y) = fraction of the data that is less than or equal to y and the cdf of the hypothesized population being sampled. The null hypothesis (ie. that a dataset has a distribution described by a fully determined (theoretical) distribution function) is rejected at a given significance level if the testing statistic exceeds the critical value.
%It is not possible to use the Kolmogorov–Smirnov test when the parameters of the distribution function are estimated from the same dataset onwhich they are tested. This, however, is usually the case in any practical application. This approach leads to type II error (failing to reject the null hypothesiswhen it is false).
%% Bemerkung: der ganze obige Absatz stammt aus: gamma-ksmirnov im Goodness-of-Fit Ordner; daher etwas umformulieren
%% this text: Kendal's Advanced Theory of Statistics, Sixth Edition, Volume 2A, by Alan Stuard et al.
%The measure of deviation used is the maximum difference between $S_n(x)$ (the cdf given by the observed sample) and $F_0(x)$ (the cdf hypothesized under $H_0$). Where $S_n(x)$ is r/n where r is the number of observations less than or equal x and n is the sample size.
%\begin{eqnarray}
%& & D_n = \underset{x}{sup}|S_n(x)-F_0(x)|
%\end{eqnarray}
%When the hypothesized distribution is composite, say $F_0(x|\theta_1,\theta_2,...\theta_s)$, with s paramters $\theta_j$, we must select s functions $t_1,...t_s$ of the $x_i$ for this purpose. [in the easy case $F_0$ is uniformly distributed on (0,1)] The transformed variables are now:
%\begin{eqnarray}
%& & y_i = \int_{-\infty}^{x_i} f_0(t_1,t_2,..t_s) du
%\end{eqnarray}
%but they are neither independent nor uniformly distributed, and their disribution depend on general both on the hypothetical distribution $F_0$ and on the true values of its parameters. So in our 3--parameter case there is some problem. 
%Perhaps the most serious limitation is that the distribution must be fully specified. That is, if location, scale, and shape parameters are estimated from the data, the critical region of the K-S test is no longer valid. It typically must be determined by simulation.
%% http://www.itl.nist.gov/div898/handbook/eda/section3/eda35g.htm
%% sup ist hier das Maximum
%% ist der Wert grösser als der kritische Wert, wird die H0 abgelehnt
%% http://de.wikipedia.org/wiki/Kolmogorow-Smirnow-Test
%% There is some more refined distribution theory for the KS test with estimated parameters (see Durbin, 1973), but that is not implemented in ks.test.
%% Durbin, J. (1973) Distribution theory for tests based on the sample distribution function. SIAM. 
%% A Kolmogorov - Smirnov goodness-of-fit test for the two-parameter weibull distribution when the parameters are estimated from the data 
%http://www.sciencedirect.com/science?_ob=ArticleURL&_udi=B6V47-4D452FR-4P&_user=10&_coverDate=12%2F31%2F1982&_rdoc=1&_fmt=high&_orig=search&_sort=d&_docanchor=&view=c&_searchStrId=1385134404&_rerunOrigin=google&_acct=C000050221&_version=1&_urlVersion=0&_userid=10&md5=3102a52af9d8937550065045f0b94ee3
%%Handbook of parametric and nonparametric statistical procedures von Sheskin David wäre nützlich
%\subsubsection{The $\chi^2$ Test}
%The main problem is to determine the number of classes k (data is grouped into k classes where we then calculate the difference between observed and expected frequencies). In a continouus distribution case, there are no natural boundaries. One procedure advocated by Mann and Gumbel was to choose the classes so that - given k - the hypothetical probabilites are all equal to 1/k. Recommendations: use classes with equal or nearly equal probabilities. Determine number of classes by:
%\begin{eqnarray}
%& & k = b (\frac{2^{1/2}(n-1)}{\lambda_{\alpha} + G^{-1}P_0})
%\end{eqnarray}
%with b between 2 and 4; where $\lambda_{\alpha}$ is the critical value (ie. 1.64 in normal distribution case); P is the value of the power function, and $G^{-1}$ comes from ????.
%In case of ML estimation, critical values should be adjusted upwards.
%The test statistic follows, approximately, a chi-square distribution with (k - c) degrees of freedom where k is the number of non-empty cells and c = the number of estimated parameters (including location and scale parameters  and shape parameters) for the distribution + 1.
%\begin{eqnarray}
%& & \chi^2 = \sum_{i=1}^k \frac{O_i - E_i}{E_i}
%\end{eqnarray}
%where the $O_i$ are the observed frequencies and the $E_i$ the expected frequencies. Null hypothesis is rejected if: $\chi^2 $>$ \chi^2_{\alpha, k-c}$.
%One rule would be to let k be equal to  $3.765(N-1)^{2/5}$ for a 5\% significance level, when using intervals with equal probability under $H_0$. But the statistical community still does not agree on the optimal choice of k. [see 15715341- goodness of fit folder, untere und obere grenze sind +- infinity]

@@

\subsection{Expected Results}

@@

\begin{table}{tbh}
\begin{center}
\begin{tabular}{llll}
\hline
Country        & Type   & \multicolumn{2}{l}{Breakpoints}  \\ \hline
Austria        & EURO   & Sep 2007 &          \\
Belgium        & EURO   & Dec 1999 &          \\
Czech Republic &        & Jul 1998 &          \\
Denmark        & ERM~II & Jun 2000 &          \\
Estonia        &        & TODO     &          \\
Finland        & EURO   & none     &          \\
France         & EURO   & Dec 2004 &          \\
Germany        & EURO   & May 2000 & Dec 2004 \\
Greece         & EURO   & none     &          \\
Hungary        &        & May 1998 &          \\
Ireland        & EURO   & Mar 2008 &          \\
Italy          & EURO   & May 1996 & Dec 2000 \\
Luxembourg     & EURO   & Dec 1998 &          \\
Netherlands    & EURO   & none     &          \\
Poland         &        & May 2001 &          \\
Portugal       & EURO   & Jul 1992 & Mar 2004 \\
Slovakia       & EURO   & Apr 1997 & Feb 2004 \\
Slovenia       & EURO   & Jul 2003 &          \\
Spain          & EURO   & May 1996 & Dec 2000 \\
Sweden         &        & Jan 1993 &          \\
United Kingdom &        & Apr 1992 &          \\ \hline
\end{tabular}
\caption{\label{tab:breakpoints} Caption needed.}
\end{center}
\end{table}

@@

\subsection{Belgium}
%% http://www.indexmundi.com/belgium/inflation_rate_%28consumer_prices%29.html
% Country Report Belgium  bei ebscohost in TX all text
There was a  very sharp fall in yearly inflation in this time (from 2\% in December 1999 to 0.2\% in January 2000). Volatility  a very clear increase in volatility; from 1999 to 2000 we had a 137\% increase in the yearly inflation rate. Economic reason: belgish franc was pegged to German Mark (since 1990). Possibly: Domestic inflation was mainly driven by the surge in world oil prices, with the result that energy prices in Belgium were around 25\% higher in the first quarter of this year than in the year-earlier period. Also a general upward trend in prices.
% source: Belgium_3206697
tax reductions in personal income tax
% source: Belgium_4611003
inflation turns out to be more volatile in small, open economies such as Belgium (might be the case in some countries, but not in all of them). some explanations happen to be: a) a more volatile currency risk premium (which cannot be the case in Belgium, since it happend after 1.1999 and the currency was pegged thereafter, and even before pegged to Germany); b) degree of exchange--rate--passthrough to inflation (not possible as well) and c) size of monetary policy shocks (if there were any, should have to be observed in ALL EMU countries, since monetary policy equal). another possible explanation are productivity shocks which boost inflation rate volatility (but this for more than 10 years????)

@@

\subsection{Italy}
Drop in yearly inflation rate towards lower mean around 1996; in December 2000 nothing really visible; in 2000 beginning of a much more volatile regime.
Economic reasons: in November 1996 Italy rejoined the european monetary system (left it in 1992); in September 1996, the new Cabinet announced Italy's decision to take part in the Monetary Union. After 2000 Italy entered an econmic crisis [auf Auffälligkeiten eingehen]
%(http://users.ox.ac.uk/~hine/seminarpapers/Visco.htm)
%in 1999 italy qualifies for EMU
%strong economic growth from 1996-2001
%2000: economic crisis

@@

\bibliographystyle{abbrvnat}
\bibliography{papers}
